---
title: "Group 3 Project"
output: html_notebook
---

```{r}
library(lehdr)
library(dplyr)
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(sf)
library(tmap)
library(tidycensus)
library(plotly)
library(ggforce)
```


```{r}
census_api_key("260b47c83abf4017c26b5022527d1558744b41cf", overwrite = TRUE)
v20 = load_variables(2018,"acs5")
```

```{r}
tmap_mode("view")

df_census=get_acs(geography = "tract",
                  variables=c("under_10000"="B19001_002", 
                              "10000_to_14999"="B19001_003", 
                              "15000_to_19999"="B19001_004", 
                              "20000_to_24999"="B19001_005", 
                              "25000_to_29999"="B19001_006", 
                              "30000_to_34999"="B19001_007", 
                              "35000_to_39999"="B19001_008", 
                              "40000_to_44999"="B19001_009", 
                              "45000_to_49999"="B19001_010", 
                              "50000_to_59999"="B19001_011", 
                              "60000_to_74999"="B19001_012", 
                              "75000_to_99999"="B19001_013",
                              "100000_to_124999"="B19001_014", 
                              "125000_to_149999"="B19001_015", 
                              "150000_to_199999"="B19001_016", 
                              "200000_or_more"="B19001_017"),
                  state="DC",geometry=TRUE,year=2018) 
df_cens=df_census %>% select(-moe) %>% spread(variable,estimate) 
```

```{r}
n_pages <- 20 
for (i in seq_len(n_pages)) {
  ggplot(data=df_census, mapping=aes(x=variable)) + geom_bar(aes(weight=estimate)) + facet_wrap_paginate(~GEOID, nrow = 3, ncol = 3, page = i)
}
```

```{r}

```


B19013_001

income in census
less than  $20,000
$20,000 to $34,999
$35,000 to $49,999
$50,000 to $74,999
$75,000 to $99,999
$100,000 to $149,999
$150,000 or more

#Do "rich" neighborhoods bike-share differently than poor?
##Game plan
- define differently, rich, and poor
- Does resident income or worker income affect bikeshare in a neighborhood more?
- Different ways to measure income. Mean 
    - Distributions of income as opposed to single value
- Look at v20 dataset
    - have some categories broken down into income brackets
    - how many people are in each bracket 
    - you can pull how many households, people... in block
        - there may be breakdown of individuals, as well as households
        - now we have essentially two histograms want to show
            - can draw histogram for any census block
            - can either plot histogram side by side
            - look at histograms, see what's there. If you see other characteristics, it will give you a perspective of approach  
- Compare visually, pull data of breakdown things then see how they are different


B19013A_001
B19013B_001
"median_inc"="B06011_001"



















