---
title: "Mob 3 Project"
output: html_notebook
---

```{r}
library(lehdr)
library(dplyr)
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(sf)
library(tmap)
library(tidycensus)
```


```{r}
census_api_key("260b47c83abf4017c26b5022527d1558744b41cf", overwrite = TRUE)
v20 = load_variables(2018,"acs5")

```

```{r}
tmap_mode("view")

df_census=get_acs(geography = "tract",
                  variables=c("median_inc"="B06011_001"),
                  state="DC",geometry=TRUE,year=2018) 
df_cens=df_census %>% select(-moe) %>% spread(variable,estimate) 

tm_shape(df_cens) +tm_polygons("median_inc",alpha=.5)
```

B19013_001

income in census
less than  $20,000
$20,000 to $34,999
$35,000 to $49,999
$50,000 to $74,999
$75,000 to $99,999
$100,000 to $149,999
$150,000 or more

#Do "rich" neighborhoods bike-share differently than poor?
##Game plan
- figure out what data to use
- define differently, rich, and poor
- Does resident income or worker income affect bikeshare in a neighborhood more?
- Different ways to measure income. Mean 
    - Distributions of income as opposed to single value
- Look at v20 dataset
    - have some categories broken down into income brackets
    - how many people are in each bracket 
    - you can pull how many households, people... in block
        - there may be breakdown of individuals, as well as households
        - now we have essentially two histograms want to show
            - can draw histogram for any census block
            - can either plot histogram side by side
            - look at histograms, see what's there. If you see other characteristics, it will give you a perspective of approach  
- Compare visually, pull data of breakdown things then see how they are different