---
title: "Group 3 Project"
output: html_notebook
---
The question proposed is the following: Do "rich" neighborhoods bike-share differently than poor?

In order to do so, census data and bikeshare data are used. To access census data, an api key is needed.
This can be obtained by visiting https://api.census.gov/data/key_signup.html 

## Loading Packages
```{r}
library(lehdr)
library(dplyr)
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(sf)
library(tmap)
library(tidycensus)
library(plotly)
library(ggforce)
library(tigris)
library(forcats)
```

## Census Data Available
```{r}
census_api_key("260b47c83abf4017c26b5022527d1558744b41cf", overwrite = TRUE)
v20 = load_variables(2018,"acs5")
```
```{r}
df_census_3=get_acs(geography = "tract",
                  variables=c("median_income" = "B19013_001"),
                  state="DC",geometry=TRUE,year=2018) 
df_cens=df_census %>% select(-moe) %>% spread(variable,estimate) 
```
```{r}
bikeshare <- read_csv(here("data_raw", "202209-capitalbikeshare-tripdata.zip"))
```
```{r}
median_inc <- df_census_3 %>% select(1,4,6)
bikeshare_start <- bikeshare %>% select(6,9,10)
```

```{r}
neigh = st_read(here("data_raw","dc_neigh.geojson")) %>% clean_names()
```
```{r}
bikeshare_start_join = st_join(bikeshare_start,neigh,largest=TRUE)
```



```{r}

```



```{r}
tmap_mode("view")

df_census=get_acs(geography = "tract",
                  variables=c("under_10000"="B19001_002", 
                              "10000_to_14999"="B19001_003", 
                              "15000_to_19999"="B19001_004", 
                              "20000_to_24999"="B19001_005", 
                              "25000_to_29999"="B19001_006", 
                              "30000_to_34999"="B19001_007", 
                              "35000_to_39999"="B19001_008", 
                              "40000_to_44999"="B19001_009", 
                              "45000_to_49999"="B19001_010", 
                              "50000_to_59999"="B19001_011", 
                              "60000_to_74999"="B19001_012", 
                              "75000_to_99999"="B19001_013",
                              "100000_to_124999"="B19001_014", 
                              "125000_to_149999"="B19001_015", 
                              "150000_to_199999"="B19001_016", 
                              "200000_or_more"="B19001_017"),
                  state="DC",geometry=TRUE,year=2018) 
df_cens=df_census %>% select(-moe) %>% spread(variable,estimate) 
```

## Reording Mapping 
We create a mapping to reorder the income brackets from lowest (under 10,000) to highest (200,000 or more). n total, there are 16 different income brackets. 
```{r}
df_census_2 = df_census %>% mutate(order = case_when(
    variable == "under_10000" ~ 1, 
    variable == "10000_to_14999" ~ 2, 
    variable == "15000_to_19999" ~ 3,
    variable == "20000_to_24999" ~ 4,
    variable == "25000_to_29999" ~ 5,
    variable == "30000_to_34999" ~ 6,
    variable == "35000_to_39999" ~ 7,
    variable == "40000_to_44999" ~ 8,
    variable == "45000_to_49999" ~ 9,
    variable == "50000_to_59999" ~ 10,
    variable == "60000_to_74999" ~ 11,
    variable == "75000_to_99999" ~ 12,
    variable == "100000_to_124999" ~ 13,
    variable == "125000_to_149999" ~ 14,
    variable == "150000_to_199999" ~ 15,
    variable == "200000_or_more" ~ 16,))
```


## Classifying Bikeshare Data into Block Groups 
```{r}
library(sf)

bikeshare_5 <- bikeshare %>% select(9,10)

ca <- tidycensus::get_acs(state = "DC", geography = "block group",
              variables = "B19013_001", geometry = TRUE)

bikeshare_5 = bikeshare_5 %>%  st_as_sf(coords = c("start_lng", "start_lat"),crs = st_crs(ca))
bikeshare_5$tract <- as.numeric(st_within(bikeshare_5, ca)) # this is fast for 1e6 points


head(bikeshare_5) # tract is the row-index for overlapping census tract record in 'ca'

# but part would take forever with 1e6 points
library(ggplot2)

ggplot(ca) +
  geom_sf() +
  geom_sf(data = bikeshare_5, aes(color = is.na(tract)))


## -------------- https://stackoverflow.com/questions/52248394/get-census-tract-from-lat-lon-using-tigris

library(tigris)
library(tidyverse)
library(sf)

ca_tracts <- tracts("DC", class = "sf") %>%
  select(GEOID, TRACTCE)

bbox <- st_bbox(ca_tracts)

my_points_tract <- st_join(bikeshare_5, ca_tracts) 
my_points_tract_revised <- na.omit(my_points_tract)

```

## Function for Visualizing Income Data
We create a function pointing to a ggplot with the arguments graph_number. This argument can take the value from 1 to 20. Each page refers to a different grouping of 9 graphs. 
```{r}
make_graphs <- function(graph_number){
    ggplot(data=df_census_2, mapping=aes(x= fct_reorder(variable, order, .desc = FALSE))) +
        geom_bar(aes(weight=estimate), fill="navyblue") + 
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
        facet_wrap_paginate(~GEOID, nrow = 3, ncol = 3, page = graph_number) + 
        xlab("Income Range, increasing") +
        ylab("Number of People")+
        ggtitle("Count per Income Bracket in DC Census Region")
}
```

## Visualizing Graph Data
We call the function on numbers 1 - 20 for 20 pages of 9 graphs each. 
```{r}
make_graphs(1)
make_graphs(2)
make_graphs(3)
make_graphs(4)
make_graphs(5)
make_graphs(6)
make_graphs(7)
make_graphs(8)
make_graphs(9)
make_graphs(10)
make_graphs(11)
make_graphs(12)
make_graphs(13)
make_graphs(14)
make_graphs(15)
make_graphs(16)
make_graphs(17)
make_graphs(18)
make_graphs(19)
make_graphs(20)
```

```{r}
# taking median of median income 
median_inc_2 <- na.omit(median_inc)
overall_median <- median(median_inc_2$estimate)

median_inc_2 <- median_inc_2 %>%
  mutate(classification = case_when(
    estimate < overall_median ~ "below_median",
    estimate == overall_median ~ "median",
    estimate > overall_median ~ "above_median"
  ))




```


B19013_001

income in census
less than  $20,000
$20,000 to $34,999
$35,000 to $49,999
$50,000 to $74,999
$75,000 to $99,999
$100,000 to $149,999
$150,000 or more

#Do "rich" neighborhoods bike-share differently than poor?
##Game plan
- define differently, rich, and poor
- Does resident income or worker income affect bikeshare in a neighborhood more?
- Different ways to measure income. Mean 
    - Distributions of income as opposed to single value
- Look at v20 dataset
    - have some categories broken down into income brackets
    - how many people are in each bracket 
    - you can pull how many households, people... in block
        - there may be breakdown of individuals, as well as households
        - now we have essentially two histograms want to show
            - can draw histogram for any census block
            - can either plot histogram side by side
            - look at histograms, see what's there. If you see other characteristics, it will give you a perspective of approach  
- Compare visually, pull data of breakdown things then see how they are different


B19013A_001
B19013B_001
"median_inc"="B06011_001"



















