---
title: "Final Mile Original"
author: "Joseph LaPell"
date: "2022-12-04"
output: html_document
---

This file is to be used for the original final mile analysis. Each person's own analysis will stem from this in new files.

Load packages
```{r}
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(sf)
library(tmap)
library(geojsonsf)
library(matrixStats)
library(geosphere)
library(lehdr)
library(tidycensus)
library(sp)
bikeshare = read_csv(here("data_raw", "202209-capitalbikeshare-tripdata.zip")) %>%
  clean_names()
metro_stations = read_csv(here("data_raw", "Metro_Stations_in_DC.csv")) %>%
  clean_names()
```

Define the Haversine distance function so we can compute distance between things
```{r}
# This function takes five arguments:
# 
# Longitude and latitude of the first location
# Longitude and latitude of the second location
# A parameter by which to round the responses

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

Just doing a sample for a Monday, this is for September 5, 2022, the first Monday of the month.
```{r}
bikeshare1 = bikeshare %>% select(ride_id, slat = start_lat, slng = start_lng) %>% na.omit() 
bikeshare1_end = bikeshare %>% select(ride_id, elat = end_lat, elng = end_lng) %>% na.omit() 
metro_stations1 = metro_stations %>% select(end_station_name = name, lat = y, lng = x) %>% na.omit()
```

https://stackoverflow.com/questions/58831578/minimum-distance-between-lat-long-across-multiple-data-frames
Finds the closest station to bikes endpoint, adds station name and distance to the bikshare dataframe.

```{r}
#calculate the distance matrix
distmatrix<-distm(bikeshare1[, c(3, 2)], metro_stations1[,c(3, 2)])
#find closest column and get distance
closest<-apply(distmatrix, 1, which.min)
bikeshare1$close_station <- as.character(metro_stations1$end_station_name[closest])
bikeshare1$distance <- apply(distmatrix, 1, min)/1000
```

Counts the amount of bikes parked within 0.1524/500ft, close enough to consider bike being used for final mile.
```{r}
final_mile = bikeshare1 %>% filter(distance < 0.1524) 
new = left_join(final_mile, bikeshare1_end, by = "ride_id")
end_points = new %>% st_as_sf(new, coords = c("elng", "elat"),crs = st_crs(4326), na.fail = FALSE, dim = "end_point")
```

```{r}
final_mile1 = new %>% mutate(start_to_end = haversine(slng, slat, elng, elat))
final_mile1_start_cnt = final_mile1 %>% filter(start_to_end <= 1.60934 & start_to_end != 0) %>% select(ride_id, close_station, distance, start_to_end, elng, elat)
```

Make all of the data frames that we'll use for analysis.
```{r}
final_mile1_grouped = final_mile1_start_cnt %>% group_by(close_station) %>% 
  summarise(count = n(), avg_distance_to_bikeshare = mean(distance), 
            avg_ride_distance = mean(start_to_end))
library(geojsonsf)
metro_stations_sf = metro_stations = geojson_sf(here("data_raw", "dc_metro_stations.geojson")) %>%
  clean_names() %>% select(close_station = name, metro_location = geometry)
final_mile1_geo = left_join(final_mile1_grouped, metro_stations_sf, by="close_station")
final_mile1_geo = st_as_sf(x = final_mile1_geo, crs = st_crs(metro_stations_sf))
final_test = final_mile1_start_cnt %>% left_join(bikeshare1_end, by = "ride_id")
```


* Analyze grouped data frame through distributions and comparing counts and average distances - Tyler.
* Visualizations: lines from start to end point for final mile rides, density distributions from above - Tyler.
* Join LODES job data with bikeshare end points that fall within final mile category - see if areas with more jobs also have more final mile rides - Joseph.
* Calculate proportion of rides within 500 ft of a station that are final mile rides - Minsum.

```{r}
census_api_key("4c7e5b613afdf7d2fd6d80005f53282c9a7c1bae")
v20 = load_variables(2019,"acs5",cache=TRUE)
df_census=get_acs(geography = "tract",
                  variables=c(
                              "total_pop"="B01003_001",
                              "worker_living"="B08301_001"),
                  state="DC",geometry=TRUE,year=2019) 
df_cens=df_census %>% rename(val=estimate) %>%  
  pivot_wider(names_from = "variable",
              values_from = c("val","moe")) %>%
  clean_names()
dc_wac <- grab_lodes(state = c("dc"), year = 2019, lodes_type = "wac", job_type = "JT00", 
           segment = "S000", state_part = "aux", agg_geo = "tract") %>% filter(state=="DC")
df_c=df_cens %>% left_join(dc_wac,by=c("geoid"="w_tract"))
```


```{r}
count_at_point = final_mile1_geo %>% select(count, metro_location)
df_c2 = df_c %>% select(total_workers = C000) %>%
  st_transform(df_c, crs = 4326)
workers_and_rides = st_join(df_c2, count_at_point)
workers_and_rides[is.na(workers_and_rides)] = 0
```

```{r}
workers_and_rides %>% select(total_workers, ride_count = count) %>%
  tm_shape() + tm_polygons(c("total_workers", "ride_count"), alpha = 0.4) + tmap_mode("view")

```

```{r}
df_ratio = workers_and_rides %>% #renames while selecting
  mutate(ratio=count/total_workers)
  df_ratio%>%
  tm_shape()+tm_polygons(c("ratio"),alpha=.4) + tmap_mode("view")
```

