---
title: "Group 3 Project"
author: "Rosalina Delwiche, Liam Durkin, Stephen Miner"
output: html_notebook
---
The question proposed is the following: Do "rich" neighborhoods bike-share differently than poor?

In order to do so, census data and bikeshare data are used. To access census data, an api key is needed.
This can be obtained by visiting https://api.census.gov/data/key_signup.html 

#Do "rich" neighborhoods bike-share differently than poor?
##Game plan - Preliminary notes
- define differently, rich, and poor
- Does resident income or worker income affect bikeshare in a neighborhood more?
- Different ways to measure income. Mean 
    - Distributions of income as opposed to single value
- Look at v20 dataset
    - have some categories broken down into income brackets
    - how many people are in each bracket 
    - you can pull how many households, people... in block
        - there may be breakdown of individuals, as well as households
        - now we have essentially two histograms want to show
            - can draw histogram for any census block
            - can either plot histogram side by side
            - look at histograms, see what's there. If you see other characteristics, it will give you a perspective of approach  
- Compare visually, pull data of breakdown things then see how they are different




## Loading Packages
```{r}
library(knitr)
library(lubridate)
library(here)
library(sf)
library(tmap)
library(forcats)
library(tigris)
library(ggforce)
library(plotly)
library(tidyverse)
library(lehdr)
library(dplyr)
library(tidycensus)
library(janitor)
```

## Access Census Data Available
```{r}
census_api_key("260b47c83abf4017c26b5022527d1558744b41cf", overwrite = TRUE)
v20 = load_variables(2018,"acs5")
```

## Extract the Median Income information for Washington DC in 2018
```{r}
df_census_3=get_acs(geography = "tract",
                  variables=c("median_income" = "B19013_001"),
                  state="DC",geometry=TRUE,year=2018) 
#df_cens=df_census %>% select(-moe) %>% spread(variable,estimate) 
```

## Import the Bikeshare data
```{r}
bikeshare <- read_csv(here("data_raw", "202209-capitalbikeshare-tripdata.zip"))
```


```{r}
median_inc <- df_census_3 %>% select(1,4,6) #Select GEOID, estimate (median inc), and geometry.
bikeshare_start <- bikeshare %>% select(6,9,10) # Select start_station_id, start_lat, start_lng
```

```{r}
neigh = st_read(here("data_raw","dc_neigh.geojson")) %>% clean_names()
```

## Get cencus data and organize it by median income ranges
```{r}
tmap_mode("view")

df_census=get_acs(geography = "tract",
                  variables=c("under_10000"="B19001_002", 
                              "10000_to_14999"="B19001_003", 
                              "15000_to_19999"="B19001_004", 
                              "20000_to_24999"="B19001_005", 
                              "25000_to_29999"="B19001_006", 
                              "30000_to_34999"="B19001_007", 
                              "35000_to_39999"="B19001_008", 
                              "40000_to_44999"="B19001_009", 
                              "45000_to_49999"="B19001_010", 
                              "50000_to_59999"="B19001_011", 
                              "60000_to_74999"="B19001_012", 
                              "75000_to_99999"="B19001_013",
                              "100000_to_124999"="B19001_014", 
                              "125000_to_149999"="B19001_015", 
                              "150000_to_199999"="B19001_016", 
                              "200000_or_more"="B19001_017"),
                  state="DC",geometry=TRUE,year=2018) 
df_cens=df_census %>% select(-moe) %>% spread(variable,estimate) 
```

## Reording Mapping 
We create a mapping to reorder the income brackets from lowest (under 10,000) to highest (200,000 or more). n total, there are 16 different income brackets. 
```{r}
df_census_2 = df_census %>% mutate(order = case_when(
    variable == "under_10000" ~ 1, 
    variable == "10000_to_14999" ~ 2, 
    variable == "15000_to_19999" ~ 3,
    variable == "20000_to_24999" ~ 4,
    variable == "25000_to_29999" ~ 5,
    variable == "30000_to_34999" ~ 6,
    variable == "35000_to_39999" ~ 7,
    variable == "40000_to_44999" ~ 8,
    variable == "45000_to_49999" ~ 9,
    variable == "50000_to_59999" ~ 10,
    variable == "60000_to_74999" ~ 11,
    variable == "75000_to_99999" ~ 12,
    variable == "100000_to_124999" ~ 13,
    variable == "125000_to_149999" ~ 14,
    variable == "150000_to_199999" ~ 15,
    variable == "200000_or_more" ~ 16,))
```


## Classifying Bikeshare Data into Block Groups + Visualization
```{r}
bikeshare_5 <- bikeshare %>% select(3,4,9,10) # Select started_at, ended_at, geometry, tract

ca <- tidycensus::get_acs(state = "DC", geography = "block group",
              variables = "B19013_001", geometry = TRUE)

bikeshare_5 = bikeshare_5 %>%  st_as_sf(coords = c("start_lng", "start_lat"),crs = st_crs(ca))
bikeshare_5$tract <- as.numeric(st_within(bikeshare_5, ca)) # this is fast for 1e6 points


head(bikeshare_5) # tract is the row-index for overlapping census tract record in 'ca'

# but part would take forever with 1e6 points

ggplot(ca) +
  geom_sf() +
  geom_sf(data = bikeshare_5, aes(color = is.na(tract)))


## -------------- https://stackoverflow.com/questions/52248394/get-census-tract-from-lat-lon-using-tigris

ca_tracts <- tracts("DC", class = "sf") %>%
  select(GEOID, TRACTCE)

bbox <- st_bbox(ca_tracts)

my_points_tract <- st_join(bikeshare_5, ca_tracts) 
my_points_tract_revised <- na.omit(my_points_tract)

```

## Function for Visualizing Income Data
We create a function pointing to a ggplot with the arguments graph_number. This argument can take the value from 1 to 20. Each page refers to a different grouping of 9 graphs. 
```{r}
make_graphs <- function(graph_number){
    ggplot(data=df_census_2, mapping=aes(x= fct_reorder(variable, order, .desc = FALSE))) +
        geom_bar(aes(weight=estimate), fill="navyblue") + 
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
        facet_wrap_paginate(~GEOID, nrow = 3, ncol = 3, page = graph_number) + 
        xlab("Income Range, increasing") +
        ylab("Number of People")+
        ggtitle("Count per Income Bracket in DC Census Region")
}
```

## Visualizing Graph Data
We call the function on numbers 1 - 20 for 20 pages of 9 graphs each. 
```{r}
make_graphs(1)
make_graphs(2)
make_graphs(3)
make_graphs(4)
make_graphs(5)
make_graphs(6)
make_graphs(7)
make_graphs(8)
make_graphs(9)
make_graphs(10)
make_graphs(11)
make_graphs(12)
make_graphs(13)
make_graphs(14)
make_graphs(15)
make_graphs(16)
make_graphs(17)
make_graphs(18)
make_graphs(19)
make_graphs(20)
```
## Combine Bikeshare and Census data together to cross reference rides with median income + Visualization
```{r}
# taking median of median income 
median_inc_2 <- na.omit(median_inc)
overall_median <- median(median_inc_2$estimate)

median_inc_2 <- median_inc_2 %>%
  mutate(classification = case_when(
    estimate < overall_median ~ "below_median",
    estimate == overall_median ~ "median",
    estimate > overall_median ~ "above_median"
  ))

ca_tracts <- median_inc %>% select(GEOID)

bbox <- st_bbox(ca_tracts)

my_points_tract <- st_join(bikeshare_5, ca_tracts)
my_points_tract_revised <- na.omit(my_points_tract)
not_analysis <- my_points_tract_revised %>% filter(!(GEOID %in% median_inc$GEOID))
unique(not_analysis$GEOID)

classifications <- as.data.frame(median_inc_2) %>% select(1,4) #select GEOID and classifications

bikeshare_with_times <- my_points_tract_revised %>%
  mutate(hour_of_day=hour(started_at))

bikeshare_with_times_2 <- left_join(bikeshare_with_times,classifications,by="GEOID") %>% na.omit(classification)

bikeshare_with_times_2 <- bikeshare_with_times_2 %>%
  group_by(classification,hour_of_day) %>%
  summarise(n=n())

ggplot(bikeshare_with_times_2, aes(x=hour_of_day)) + geom_bar(aes(weight=n)) + facet_wrap(~classification, scale="free")
```

```{r}
median_inc_ranked <- median_inc_2
median_inc_ranked$rank = rank(desc(median_inc_ranked$estimate))

median_inc_subset <- median_inc_ranked %>% slice(which(rank>=1 & rank<=10 | rank<=177 & rank>=166))

classifications_2 <- as.data.frame(median_inc_subset) %>% select(1,3)

bikeshare_with_times_3 <- left_join(bikeshare_with_times,classifications_2,by="GEOID") %>% na.omit(classification)

bikeshare_with_times_3 <- bikeshare_with_times_3 %>%
  group_by(classification,hour_of_day) %>%
  summarise(n=n())
ggplot(bikeshare_with_times_3, aes(x=hour_of_day)) + geom_bar(aes(weight=n)) + facet_wrap(~classification, scale="free")
```
```{r}
bikeshare_durations <- my_points_tract_revised %>%
  mutate(duration=difftime(ended_at,started_at,units="mins"))
classifications_3 <- as.data.frame(median_inc_2) %>% select(1,4)
bikeshare_durations <- left_join(bikeshare_durations,classifications_3,by="GEOID") %>% na.omit(classification)
bikeshare_durations <- bikeshare_durations %>%
  group_by(classification,duration) %>%
  summarise(n=n())
bikeshare_durations <- bikeshare_durations %>% filter(classification != "median")
ggplot(bikeshare_durations, aes(x=duration)) + geom_bar(aes(weight=n)) + facet_wrap(~classification, scale="free")
```
## Duration Analysis
Analyze the duration of the bike ride against the neighborhoods above the median income and below the median income
```{r}
bikeshare_durations <- bikeshare_durations %>% filter(duration<=250, classification != "median")
ggplot(bikeshare_durations, aes(x=duration)) + geom_bar(aes(weight=n)) + facet_wrap(~classification, scale="free")
```
There are some very long bike rides, so the scope is narrowed down to a duration of less than 100 minutes
```{r}
bikeshare_durations <- bikeshare_durations %>% filter(duration<=100, classification != "median")
ggplot(bikeshare_durations, aes(x=duration)) + geom_bar(aes(weight=n)) + facet_wrap(~classification, scale="free")
```
We can look to increase our view further by changing the duration to be less than or equal to 50
```{r}
bikeshare_durations <- bikeshare_durations %>% filter(duration<=50, classification != "median")
ggplot(bikeshare_durations, aes(x=duration)) + geom_bar(aes(weight=n)) + facet_wrap(~classification, scale="free_y")
```

Conclusion:  It seems that the above median group uses the bikeshare program more than the other grous at or below the median
